---
title: 'Spatiotemporal analysis of cervical cancer incidence among HIV positive women in South Africa: Type II-IV interactions'
author: "Garyfallos Konstantinoudis"
output:
  html_document:
    toc: true
    toc_float: true
bibliography: biblio.bib
---


<style type="text/css">
body{ /* Normal  */
      font-size: 14px;
  }
h1.title {
  font-size: 30px;
  color: black;
  font-weight: bold;
}
h1 { /* Header 1 */
    font-size: 25px;
  color: black;
  font-weight: bold;
}
h2 { /* Header 2 */
    font-size: 20px;
  color: black;
  font-weight: bold;
}
h3 { /* Header 3 */
    font-size: 15px;
  color: black;
  font-weight: bold;
}
code.r{ /* Code block */
    font-size: 14px;
}
pre, code {
    color: 	#1B0F0E;
  }
</style>


\pagenumbering{gobble} 
\pagenumbering{arabic} 



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE, fig.align = "center",
                      fig.width=7, fig.height=4)
```


In this practical you will use `R` [@R-base] as well as `nimble` [@nimble] to carry out a spatial and a spatio-temporal small area disease risk analysis.

In particular, you are going to model yearly cervical cancer (C53.0, C53.1, C53.8 and C53.9) incidence among women diagnosed with HIV in South Africa during 2004-2014. This dataset was previously explored in the tutorial entitled "Spatiotemporal analysis of Cervical cancer among HIV positive women in South Africa: type I spatiotemporal interaction".

The code used for this tutorial is an adaptation of the code provided in Chapter 15 in [@haining2020].

## Before starting the practical

* Load needed libraries:
```{r echo=TRUE, eval = TRUE, message=FALSE}
library(dplyr)        # A package for data manipulation
library(sf)           # Simple feature for R
library(spdep)        # Functions and tests for evaluating spatial patterns 
                      # and autocorrelation
library(tidyr)

library(nimble)       # A package for running nimble from R
library(ggmcmc)       # A package for plotting and viewing of MCMC output
library(coda)         # A package for summarizing and plotting of MCMC output 
                      # and diagnostic tests

library(ggplot2)      # A package that implements the grammar of graphics, which is a term used to
                      # break up graphs into semantic components, such as geometries and layers.
library(viridis)      # A package providing color palettes 
library(patchwork)

# For tables in RMarkdown
library(knitr)
library(kableExtra)
```

* To install the entire suite of packages, we can use:
```{r eval = FALSE,  results="hide"}

install.packages(c("sf", "dplyr", "tidyr", "nimble", "coda", "spdep", "patchwork", "ggplot2","GGally", "ggmcmc", "viridis", "patchwork", "knitr", "kableExtra"),
                 dependencies = TRUE, repos = "http://cran.r-project.org")

```

## Data

1. Import the`.rds` file with the data and call the data.frame object as *CC_HIV*. Import also the `.shp` and call it `RSA_shp`.

```{r eval=TRUE, echo=TRUE, message=FALSE}
CC_HIV <- readRDS("CC_HIV")
head(CC_HIV)

RSA_shp <- read_sf("ShapeSA_mun.shp")
```
Here The first column labeled `year` is the year of the diagnosis, `ID` is a municipality ID, `CCcases` is the number of cervical cancer cases and `HIV_cases_expected` is the number of expected number of cases (women) diagnosed with HIV.

## Space time interaction type II

7. Write a model in `nimble` were you include spatially and temporally structured and unstructured components.

```{r}

st_typeII_BYMmodel <- nimbleCode(
{
  for (i in 1:N) 
    {
    for(t in 1:Temporal)
      {
          
      O[i,t]  ~ dpois(mu[i,t])
      log(mu[i,t]) <- log(E[i,t]) + alpha + theta[i] + phi[i] + gamma[t] + xi[t] + zeta[i, t]
      
      gamma[t] ~ dnorm(0,tau.gamma) 
    }
    
      zeta[i, 1:Temporal] ~ dcar_normal(adj.tm[1:K], weights.tm[1:K], 
                                        num.tm[1:Temporal], tau.zeta, zero_mean = 1)
      theta[i] ~ dnorm(0,tau.theta)                 
    }
  

# intrinsic CAR prior on temporal random effects
  xi[1:Temporal] ~ dcar_normal(adj.tm[1:K], weights.tm[1:K], 
                               num.tm[1:Temporal], tau.xi, zero_mean = 1)
  
# intrinsic CAR prior on spatial random effects
  phi[1:N] ~ dcar_normal(adj[1:L], weights[1:L], num[1:N], tau.phi, zero_mean = 1)
  

# priors
  alpha  ~ dflat()                    
  overallRR <- exp(alpha)             
    
  # temporal field
  tau.gamma~ dgamma(1,0.001)          
  sigma2.gamma <- 1/tau.gamma         
  
  tau.xi ~ dgamma(0.5,0.005)          
  sigma2.xi <- 1/tau.xi               
  
  # spatial field
  tau.theta ~ dgamma(1,0.001)         
  sigma2.theta <- 1/tau.theta         
  
  tau.phi ~ dgamma(0.5,0.005)         
  sigma2.phi <- 1/tau.phi             
  
  # spatiotemporal field
  tau.zeta ~ dgamma(1,0.001)          
  sigma2.zeta <- 1/tau.zeta      

}
)

```

8. Define the spatial weight matrix

```{r}

RSA_nb <- poly2nb(RSA_shp)
nbWB <- nb2WB(nb = RSA_nb)

```

9. Define the temporal weight matrix

```{r eval=TRUE, echo=TRUE, include=TRUE}

# For the random walk 1 we can do something similar
Temporal <- length(unique(CC_HIV$year))
W <- matrix(0, nrow = Temporal, ncol = Temporal)

for(i in 1:(Temporal-1)) W[i,i+1] <- 1
for(i in 1:(Temporal-1)) W[i+1,i] <- 1

Wnb_temporal <- mat2listw(W)
Wnb_temporal <- nb2WB(nb = Wnb_temporal$neighbours)

```

10. Create matrices of the observed and expected number of cervical cancers where each column is a different year and each row a different municipality. 

```{r}

# prepare observed data
Observed <- spread(CC_HIV[,c("year", "ID", "CCcases")], year, CCcases)
Observed <- as.matrix(Observed[,-1])

# prepare expected data
Expected <- spread(CC_HIV[,c("year", "ID", "HIV_expected")], year, HIV_expected)
Expected <- as.matrix(Expected[,-1])

```

11. Set the data and the constants.

```{r eval=TRUE, echo=TRUE, include=TRUE}

N <- nrow(Observed)

CCdata = list(
                 O = Observed                           
)      
      
CCConsts <-list(      
                 N = N,                                  
                 # space
                 L = length(nbWB$weights),               
                 E = Expected,                           
                 adj = nbWB$adj,                         
                 num = nbWB$num, 
                 weights = nbWB$weights, 
                 # time
                 Temporal = Temporal, 
                 K = length(Wnb_temporal$weights), 
                 adj.tm = Wnb_temporal$adj, 
                 num.tm = Wnb_temporal$num, 
                 weights.tm = Wnb_temporal$weights
)

```

12. Define the initial values.

```{r eval=TRUE, echo=TRUE}

inits <- list(
  list(alpha = 1,
       tau.theta = 10,
       tau.phi = 1,
       tau.xi = 8, 
       tau.gamma = 5,
       tau.zeta = 5,
       theta = rep(0.02, times = N), 
       phi = c(rep(0.2, times = N)), 
       gamma = rep(0.02, times = Temporal),
       xi = c(rep(0.2, times = Temporal)), 
       zeta = matrix(0.2, nrow = N, ncol = Temporal)
       ),
  list(alpha = 0.5,
       tau.theta = 1,
       tau.phi = 0.1,
       tau.xi = 0.8, 
       tau.gamma = 0.5,
       tau.zeta = 1, 
       theta = rep(0.05, times = N),
       phi = c(rep(-0.05, times = N)), 
       gamma = rep(-0.02, times = Temporal),
       xi = c(rep(-0.2, times = Temporal)), 
       zeta = matrix(-0.2, nrow = N, ncol = Temporal)
       )
)

```

13. Define the monitors. Monitor the overall relative risk, all the variance hyperparameters and all the random effects.

```{r eval=TRUE, echo=TRUE}
params <- c("overallRR", "sigma2.phi",
            "sigma2.theta", "phi", "theta", "xi", "gamma",
            "sigma2.gamma", "sigma2.xi", "mu",
            "sigma2.zeta", "zeta", "alpha")
```

14. Specify the MCMC setting

```{r eval=TRUE, echo=TRUE, include=TRUE}
# MCMC setting
ni <- 50000  # nb iterations 
nt <- 5      # thinning interval
nb <- 10000  # nb iterations as burn-in 
nc <- 2      # nb chains
```

15. Call the `nimleMCMC()` function. (Takes 15 minutes)

```{r eval=FALSE, echo=TRUE}

t_0 <- Sys.time()
st_typeII_BYM.model <- nimbleMCMC(code = st_typeII_BYMmodel,
                      data = CCdata,
                      constants = CCConsts, 
                      inits = inits,
                      monitors = params,
                      niter = ni,
                      nburnin = nb,
                      thin = nt, 
                      nchains = nc, 
                      setSeed = 9, 
                      progressBar = TRUE, 
                      samplesAsCodaMCMC = TRUE, 
                      summary = TRUE, 
                      WAIC = TRUE
                      )
t_1 <- Sys.time()
t_1 - t_0

```

```{r eval=TRUE, echo=FALSE, include=FALSE}

# saveRDS(st_typeII_BYM.model, file = "st_typeII_BYM.model")
st_typeII_BYM.model <- readRDS("st_typeII_BYM.model")

```


### Convergence diagnostics


16. Plot the traceplots for the intercept and the four variances

```{r eval=TRUE, echo=TRUE, fig.height = 12, fig.width=10}

ggs_stBYM.model <- ggs(st_typeII_BYM.model$samples)


ggs_stBYM.model %>% filter(Parameter %in% c("overallRR", "sigma2.theta","sigma2.phi",
                                            "sigma2.gamma", "sigma2.xi", "sigma2.zeta")) %>% 
  ggs_traceplot() + theme_bw()

```

17. Calculate the Gelman and Rubin diagnostic and check the convergence of the random effects. 

```{r eval=FALSE, echo=TRUE}

GR.diag <- gelman.diag(st_typeII_BYM.model$samples, multivariate = FALSE)

all(GR.diag$psrf[startsWith(rownames(GR.diag$psrf), c("gamma")),"Point est."] < 1.1)
all(GR.diag$psrf[startsWith(rownames(GR.diag$psrf), c("theta")),"Point est."] <1.1)
all(GR.diag$psrf[startsWith(rownames(GR.diag$psrf), c("phi")),"Point est."] < 1.1)
all(GR.diag$psrf[startsWith(rownames(GR.diag$psrf), c("xi")),"Point est."] <1.1)
all(GR.diag$psrf[startsWith(rownames(GR.diag$psrf), c("zeta")),"Point est."]<1.1)

```



## Space time interaction type III

18. Write the model for spacetime interaction type III

```{r}

st_typeIII_BYMmodel <- nimbleCode(
{
  for (i in 1:N) 
    {
    for(t in 1:Temporal)
    {
      O[i,t]  ~ dpois(mu[i,t])
      log(mu[i,t]) <- log(E[i,t]) + alpha + theta[i] + phi[i] + gamma[t] + xi[t] + zeta[i, t]
      
      zeta[i, t] <- zeta.transposed[t, i]
      zeta.transposed[t, 1:N] ~ dcar_normal(adj[1:L], weights[1:L], 
                                           num[1:N], tau.zeta, zero_mean = 1)
      
      gamma[t] ~ dnorm(0,tau.gamma) 
    }
    
      theta[i] ~ dnorm(0,tau.theta)                 
    }
  

# intrinsic CAR prior on temporal random effects
  xi[1:Temporal] ~ dcar_normal(adj.tm[1:K], weights.tm[1:K], 
                               num.tm[1:Temporal], tau.xi, zero_mean = 1)
  
# intrinsic CAR prior on spatial random effects
  phi[1:N] ~ dcar_normal(adj[1:L], weights[1:L], num[1:N], tau.phi, zero_mean = 1)
  

# priors
  alpha  ~ dflat()                    
  overallRR <- exp(alpha)             
    
  # temporal field
  tau.gamma~ dgamma(1,0.001)          
  sigma2.gamma <- 1/tau.gamma         
  
  tau.xi ~ dgamma(0.5,0.005)          
  sigma2.xi <- 1/tau.xi               
  
  # spatial field
  tau.theta ~ dgamma(1,0.001)         
  sigma2.theta <- 1/tau.theta         
  
  tau.phi ~ dgamma(0.5,0.005)         
  sigma2.phi <- 1/tau.phi             
  
  # spatiotemporal field
  tau.zeta ~ dgamma(1,0.001)          
  sigma2.zeta <- 1/tau.zeta      

}
)

```


19. Redefine the initials ones to consider the `zeta.transpose`

```{r eval=TRUE, echo=TRUE}

inits <- list(
  list(alpha = 1,
       tau.theta = 10,
       tau.phi = 1,
       tau.xi = 8, 
       tau.gamma = 5,
       tau.zeta = 5,
       theta = rep(0.02, times = N), 
       phi = c(rep(0.2, times = N)), 
       gamma = rep(0.02, times = Temporal),
       xi = c(rep(0.2, times = Temporal)), 
       zeta.transposed = t(matrix(0.2, nrow = N, ncol = Temporal))
       ),
  list(alpha = 0.5,
       tau.theta = 1,
       tau.phi = 0.1,
       tau.xi = 0.8, 
       tau.gamma = 0.5,
       tau.zeta = 1, 
       theta = rep(0.05, times = N),
       phi = c(rep(-0.05, times = N)), 
       gamma = rep(-0.02, times = Temporal),
       xi = c(rep(-0.2, times = Temporal)), 
       zeta.transposed = t(matrix(-0.2, nrow = N, ncol = Temporal))
       )
)

```


20. Call the `nimble` function (18 minutes):

```{r eval=FALSE, echo=TRUE}

t_0 <- Sys.time()
st_typeIII_BYM.model <- nimbleMCMC(code = st_typeIII_BYMmodel,
                      data = CCdata,
                      constants = CCConsts, 
                      inits = inits,
                      monitors = params,
                      niter = ni,
                      nburnin = nb,
                      thin = nt, 
                      nchains = nc, 
                      setSeed = 9, 
                      progressBar = TRUE, 
                      samplesAsCodaMCMC = TRUE, 
                      summary = TRUE, 
                      WAIC = TRUE
                      )
t_1 <- Sys.time()
t_1 - t_0

```

```{r eval=TRUE, echo=FALSE, include=FALSE}

# saveRDS(st_typeIII_BYM.model, file = "st_typeIII_BYM.model")
st_typeIII_BYM.model <- readRDS("st_typeIII_BYM.model")

```

### Convergence diagnostics

21. Plot the traceplots for the intercept and the four variances

```{r eval=TRUE, echo=TRUE, fig.height = 12, fig.width=10}

ggs_stBYM.model <- ggs(st_typeIII_BYM.model$samples)


ggs_stBYM.model %>% filter(Parameter %in% c("overallRR", "sigma2.theta","sigma2.phi",
                                            "sigma2.gamma", "sigma2.xi", "sigma2.zeta")) %>% 
  ggs_traceplot() + theme_bw()

```


21. Calculate the Gelman and Rubin diagnostic and check the convergence of the random effects. 

```{r eval=FALSE, echo=TRUE}

GR.diag <- gelman.diag(st_typeIII_BYM.model$samples, multivariate = FALSE)

all(GR.diag$psrf[startsWith(rownames(GR.diag$psrf), c("gamma")),"Point est."] < 1.1)
all(GR.diag$psrf[startsWith(rownames(GR.diag$psrf), c("theta")),"Point est."] <1.1)
all(GR.diag$psrf[startsWith(rownames(GR.diag$psrf), c("phi")),"Point est."] < 1.1)
all(GR.diag$psrf[startsWith(rownames(GR.diag$psrf), c("xi")),"Point est."] <1.1)
all(GR.diag$psrf[startsWith(rownames(GR.diag$psrf), c("zeta")),"Point est."]<1.1)

```

## Space time interaction type IV

22. Write the model for spacetime interaction type IV

```{r}

st_typeIV_BYMmodel <- nimbleCode(
{
  for (i in 1:N) 
    {
    for(t in 1:Temporal)
    {
      O[i,t]  ~ dpois(mu[i,t])
      log(mu[i,t]) <- log(E[i,t]) + alpha + theta[i] + phi[i] + 
        gamma[t] + xi[t] + zeta.centred[i, t]
      
      gamma[t] ~ dnorm(0,tau.gamma) 
      
      d[i,t] <- zeta[zetaID[i, t]]
      zeta.centred[i, t] <- d[i, t] - row.mean[i] - col.mean[t]
      zetaID[i, t] <- t + (i - 1)*Temporal
    }
    
      theta[i] ~ dnorm(0,tau.theta)                 
    }
  
  
  for (i in 1:N) {
    row.mean[i] <- sum(d[i, 1:Temporal])/Temporal
  }
  for (t in 1:Temporal) {
    col.mean[t] <- sum(d[1:N, t])/N
  }

  
# intrinsic CAR prior on temporal random effects
  xi[1:Temporal] ~ dcar_normal(adj.tm[1:K], weights.tm[1:K], 
                               num.tm[1:Temporal], tau.xi, zero_mean = 1)
  
# intrinsic CAR prior on spatial random effects
  phi[1:N] ~ dcar_normal(adj[1:L], weights[1:L], 
                         num[1:N], tau.phi, zero_mean = 1)
  
# Type IV spacetime interaction
  zeta[1:NT] ~ dcar_normal(adj.st[1:M], weights.st[1:M], 
                          num.st[1:NT], tau.zeta)

# priors
  alpha  ~ dflat()                    
  overallRR <- exp(alpha)             
    
  # temporal field
  tau.gamma~ dgamma(1,0.001)          
  sigma2.gamma <- 1/tau.gamma         
  
  tau.xi ~ dgamma(0.5,0.005)          
  sigma2.xi <- 1/tau.xi               
  
  # spatial field
  tau.theta ~ dgamma(1,0.001)         
  sigma2.theta <- 1/tau.theta         
  
  tau.phi ~ dgamma(0.5,0.005)         
  sigma2.phi <- 1/tau.phi             
  
  # spatiotemporal field
  tau.zeta ~ dgamma(1,0.001)          
  sigma2.zeta <- 1/tau.zeta      
}
)

```


23. Define the weights based on the Kronecker product. This code is available [here](https://www.sptmbook.com/datacode.html).

```{r}

contruct_typeIV_weights <- function(spW,ntimes) {
  ####    this function constructs the weights for the space-time interaction of Type IV
  ####    On input:
  ####       - spW is an NxN spatial weights matrix with diagonals=0
  ####       - ntimes denotes the number of time points in the time series
  Dsp <- apply(spW,1,sum)
  Qsp <- diag(Dsp) - spW
  Dtm <- c(1,rep(2,ntimes-2),1)
  Qtm <- diag(Dtm)
  for (tt in 1:ntimes) {
    if (tt==1) Qtm[tt,2] <- -1
    if (tt==ntimes) Qtm[tt,tt-1] <- -1
    if (tt>1 & tt<ntimes) Qtm[tt,c(tt-1,tt+1)] <- -1
  }
  QtypeIV <- kronecker(Qsp,Qtm)
  DtypeIV <- diag(diag(QtypeIV))
  WtypeIV <- DtypeIV - QtypeIV
  adj <- list(num=NULL,weights=NULL,adj=NULL)
  n <- nrow(WtypeIV)
  for (i in 1:n) {
    ids <- which(WtypeIV[i,]!=0)
    adj$num <- c(adj$num,length(ids))
    adj$adj <- c(adj$adj,ids)
    adj$weights <- c(adj$weights,WtypeIV[i,ids])
  }
  return(adj)
}


spW <- nb2mat(RSA_nb, style='B')
kronecker_weights <- contruct_typeIV_weights(spW = spW, ntimes = Temporal)

```

24. Add the Kronecker product in the constants

```{r}

CCConsts <-list(      
                 N = N,                                  
                 # space
                 L = length(nbWB$weights),               
                 E = Expected,                           
                 adj = nbWB$adj,                         
                 num = nbWB$num, 
                 weights = nbWB$weights, 
                 # time
                 Temporal = Temporal, 
                 K = length(Wnb_temporal$weights), 
                 adj.tm = Wnb_temporal$adj, 
                 num.tm = Wnb_temporal$num, 
                 weights.tm = Wnb_temporal$weights,
                 # Kronecker
                 NT = N*Temporal,
                 M = length(kronecker_weights$weights),
                 adj.st = kronecker_weights$adj, 
                 num.st = kronecker_weights$num, 
                 weights.st = kronecker_weights$weights
                 
)

```


25. Redefine the initials and monitors

```{r eval=TRUE, echo=TRUE}

inits <- list(
  list(alpha = 1,
       tau.theta = 10,
       tau.phi = 1,
       tau.xi = 8, 
       tau.gamma = 5,
       tau.zeta = 5,
       theta = rep(0.02, times = N), 
       phi = c(rep(0.2, times = N)), 
       gamma = rep(0.02, times = Temporal),
       xi = c(rep(0.2, times = Temporal)), 
       zeta = rep(0.2, times = N*Temporal)
       ),
  list(alpha = 0.5,
       tau.theta = 1,
       tau.phi = 0.1,
       tau.xi = 0.8, 
       tau.gamma = 0.5,
       tau.zeta = 1, 
       theta = rep(0.05, times = N),
       phi = c(rep(-0.05, times = N)), 
       gamma = rep(-0.02, times = Temporal),
       xi = c(rep(-0.2, times = Temporal)), 
       zeta = rep(-0.2, times = N*Temporal)
       )
)

```

```{r eval=TRUE, echo=TRUE}
params <- c("overallRR", "sigma2.phi",
            "sigma2.theta", "phi", "theta", "xi", "gamma",
            "sigma2.gamma", "sigma2.xi", "mu",
            "sigma2.zeta", "zeta", "alpha", "zeta.centred")
```

26. Call the `nimbleMCMC` function. (1.6 days)

```{r eval=FALSE, echo=TRUE}

t_0 <- Sys.time()
st_typeIV_BYM.model <- nimbleMCMC(code = st_typeIV_BYMmodel,
                      data = CCdata,
                      constants = CCConsts, 
                      inits = inits,
                      monitors = params,
                      niter = ni,
                      nburnin = nb,
                      thin = nt, 
                      nchains = nc, 
                      setSeed = 9, 
                      progressBar = TRUE, 
                      samplesAsCodaMCMC = TRUE, 
                      summary = TRUE, 
                      WAIC = TRUE
                      )
t_1 <- Sys.time()
t_1 - t_0

```

```{r eval=TRUE, echo=FALSE, include=FALSE}

# saveRDS(st_typeIV_BYM.model, file = "st_typeIV_BYM.model")
st_typeIV_BYM.model <- readRDS("st_typeIV_BYM.model")

```


### Convergence diagnostics

27. Plot the traceplots for the intercept and the four variances

```{r eval=TRUE, echo=TRUE, fig.height = 12, fig.width=10}

ggs_stBYM.model <- ggs(st_typeIV_BYM.model$samples)


ggs_stBYM.model %>% filter(Parameter %in% c("overallRR", "sigma2.theta","sigma2.phi",
                                            "sigma2.gamma", "sigma2.xi", "sigma2.zeta")) %>% 
  ggs_traceplot() + theme_bw()

```


28. Calculate the Gelman and Rubin diagnostic and check the convergence of the random effects. 

```{r eval=FALSE, echo=TRUE}

GR.diag <- gelman.diag(st_typeIV_BYM.model$samples, multivariate = FALSE)

all(GR.diag$psrf[startsWith(rownames(GR.diag$psrf), c("gamma")),"Point est."] < 1.1)
all(GR.diag$psrf[startsWith(rownames(GR.diag$psrf), c("theta")),"Point est."] <1.1)
all(GR.diag$psrf[startsWith(rownames(GR.diag$psrf), c("phi")),"Point est."] < 1.1)
all(GR.diag$psrf[startsWith(rownames(GR.diag$psrf), c("xi")),"Point est."] <1.1)
all(GR.diag$psrf[startsWith(rownames(GR.diag$psrf), c("zeta.centred")),"Point est."]<1.1)

```

## Model comparison

At this point, we will load the spacetime interaction type I shown in a previous tutorial so we can compare all 4 types of interactions.

```{r eval=TRUE, echo=FALSE, include=FALSE}

st_typeI_BYM.model <- readRDS("st_typeI_BYM.model")

```

29. Spaghetti plots

```{r, fig.height=9, fig.width=6}
dat.stre <- data.frame(
  year = rep(CC_HIV$year, times = 4),
  ID = rep(CC_HIV$ID, times = 4), 
  STRR = c(
   exp(
    st_typeI_BYM.model$summary$all.chains[startsWith(rownames(st_typeI_BYM.model$summary$all.chains), 
          "zeta"),"Median"]), 
    exp(
    st_typeII_BYM.model$summary$all.chains[startsWith(rownames(st_typeII_BYM.model$summary$all.chains), 
          "zeta"),"Median"]),
    exp(
    st_typeIII_BYM.model$summary$all.chains[startsWith(rownames(st_typeIII_BYM.model$summary$all.chains),
          "zeta"),"Median"]),
    exp(st_typeIV_BYM.model$summary$all.chains[startsWith(rownames(st_typeIV_BYM.model$summary$all.chains), 
                                                    "zeta.centred"),"Median"])
  ), 
  model = c(rep("TypeI", times = 1859), rep("TypeII", times = 1859), rep("TypeIII", times = 1859),
            rep("TypeIV", times = 1859))
)

# a. The spaghetti plot
ggplot() + geom_line(data = dat.stre, aes(x = year, y = STRR, group = ID, col = ID)) +
  scale_color_viridis_c() + theme_bw() + facet_grid(rows = vars(model))
```

30. Spatiotemporal maps

```{r, fig.height=15, fig.width=20}

dat.stre$year <- as.numeric(dat.stre$year)
dat.plot <- left_join(RSA_shp, dat.stre, by = c("ID" = "ID"))

spatimeplot <- function(j){
  
dat.plot %>% filter(model == j) %>% 
  mutate(STRRq = cut(STRR, 
                         breaks = c(0, 0.8, 1, 1.2, 1.5, 3), 
                         label = c("0.8<", "(0.8, 1]", "(1, 1.2]", "(1.2, 1.5]", ">1.5"), 
                         include.lowest = TRUE)
             ) %>% 
ggplot() + geom_sf(aes(fill = STRRq), col = NA) + facet_wrap(~year) +
scale_fill_viridis_d(name = "") + theme_bw() + 
theme(text = element_text(size = 30)) + ggtitle(j) -> p

return(p)
}

spatimeplot("TypeI")
spatimeplot("TypeII") 
spatimeplot("TypeIII")
spatimeplot("TypeIV") 

```

31. Table of hyperparameters


```{r eval=TRUE, echo=TRUE}

gethyper <- function(metric, model){
  c(
    model$summary$all.chains[paste0("sigma2.theta"),metric], 
    model$summary$all.chains[paste0("sigma2.phi"),metric], 
    model$summary$all.chains[paste0("sigma2.gamma"),metric], 
    model$summary$all.chains[paste0("sigma2.xi"),metric],
    model$summary$all.chains[paste0("sigma2.zeta"),metric]
    )
}

tabhyper <- function(model){
    dat.hyper <- 
      round(
      data.frame(
        median = 
          gethyper("Median", model),
        LL = 
          gethyper("95%CI_low", model),
        UL = 
          gethyper("95%CI_upp", model) 
        ),
      digits = 3)
    
    dat.hyper$CrI <-paste0(dat.hyper$median, " (", dat.hyper$LL, ", ", dat.hyper$UL, ")")
    return(dat.hyper[,"CrI"])
}


dat.hyper <- 
data.frame(
"Type I" = tabhyper(st_typeI_BYM.model), 
"Type II" = tabhyper(st_typeII_BYM.model),
"Type III" = tabhyper(st_typeIII_BYM.model),
"Type IV" = tabhyper(st_typeIV_BYM.model)
)



row.names(dat.hyper) <- 
      c("sigma2.theta", "sigma2.phi", "sigma2.gamma", "sigma2.xi", "sigma2.zeta")

knitr::kable(dat.hyper, caption = "Posterior median and 95% CrI of the variance hyperparameters.",
             align=c(rep('r',times=4))) %>%  
  kable_styling(bootstrap_options = "striped", full_width = F, position = "center")

```

32. WAIC
```{r}

dat.WAIC <- 
data.frame(
"Type I" = st_typeI_BYM.model$WAIC, 
"Type II" = st_typeII_BYM.model$WAIC,
"Type III" = st_typeIII_BYM.model$WAIC,
"Type IV" = st_typeIV_BYM.model$WAIC
)

knitr::kable(dat.WAIC) %>%  kable_styling(bootstrap_options = "striped", full_width = F, position = "center")

```


## Conclusion

This tutorial shows how to fit higher order interactions in `nimble`. Note that most results for the hyperparameters have not yet converged, thus the MCMC should be tuned accordingly if these parameters are of the main interest. The WAIC suggest that the best fitting model is the one with type II interaction. Nevertheless the difference between type II and type I and type III is negligible (wrt the WAIC). 

## References
